\documentclass[twocolumn,numbook]{svjour3}
\usepackage{style}
\usepackage{subfiles}
\usepackage{microtype}
\usepackage[capitalise]{cleveref}

% Title
\title{Multiset-trie data structure}
\author{Mikita Akulich \and Iztok Savnik \and Matja\v z Krnc \and Riste \v Skrekovski}
\institute{
		Mikita Akulich 
		\at 
		University of Primorska, FAMNIT, Koper, Slovenia\\
		\email{mikita.akulich@gmail.com}
		\and
		Iztok Savnik
		\at 
		University of Primorska, FAMNIT, Koper, Slovenia\\
		\email{iztok.savnik@famnit.upr.si}
		\and 
		Matja\v z Krnc 
		\at 
		University of Primorska, FAMNIT, Koper, Slovenia\\
		\email{matjaz.krnc@famnit.upr.si}
		\and 
		Riste \v Skrekovski
		\at
		University of Ljubljana, FMF, Ljubljana, Slovenia\\
		\email{riste.skrekovski@famnit.upr.si}
		}

\begin{document}
\maketitle

%Abstract
\begin{abstract}
This paper proposes a new data structure \emph{multiset-trie} that is designed for storing and efficiently processing of a set of multisets. Moreover, multiset-trie can operate on a set of sets without efficiency loss.
%
The multiset-trie is a search tree with properties similar to those of a trie. It implements all standard search tree operations together with the multiset containment operations such as sub-multiset and super-multiset. Suppose we have a set of multisets $S$ and a multiset $X$. The multiset containment operations retrieve multisets from $S$ that are either sub-multisets or super-multisets of $X$. 

We present the mathematical analysis of a multiset-trie that gives the time complexity of the algorithms and the space complexity of the data structure. Further, the empirical analysis of the data structure is implemented in a series of experiments. The experiments illuminate the time complexity space of the multiset containment operations. Finally, the comparison of the multiset-trie with the inverted index shows that the multiset-trie outperforms the inverted index by up to 4 orders of magnitude.
%
\end{abstract}
\keywords{multiset, index data structure, containment operations, inverted file}
%\subclass{68P05, 68P20, 68Q87}

% section 1
\input{tex/s_intro}
% section 2
\input{tex/s_data_structure}
% section 3
\input{tex/s_operations}
% section 4
\input{tex/s_math}
% section 5
\input{tex/s_experiments}
% section 6
\input{tex/s_relwork}
% section 7
\input{tex/s_conclusions}

\bibliographystyle{abbrv}
\bibliography{multiset-trie}

\end{document}